---
title: "An Exploratory Analysis of Police Acquisition of Militarized Equipment"
output:
  html_notebook: default
  html_document: default
---


## Setup
Load libraries and import the police acquisitions data set.
```{r, results='hide'}
library(data.table)
library(ggplot2)
library(scales)
library(ggmap)
library(statebins)
library(plyr)
library(readxl)
library(stringr)
library(stringi)
library(extrafont)
library(USA.county.data)
library(lattice)
#library(devtools)
#devtools::install_github("deleetdk/USA.county.data") # only need to load this once

data = fread("~/projects/policeMilitarySurplus/data/policeAcquisitionOfMilitarySurplus2014.csv")
colnames(data) = make.names(names(data)) # make column names R-friendly (remove spaces)
data$Item.Name = sapply(data$Item.Name, trimws) # trim whitespace
data = data[!(data$Item.Name == ""), ] # get rid of empty string value

# make an appropriate theme for our plots later on
military_theme = theme_classic() + 
  theme(plot.title=element_text(size=20, color="#0D0A0B", vjust=1, hjust=0.5, family="Courier New"),
  text=element_text(size=13, hjust=0.5, family="Courier New"), plot.background=element_rect(color="#F3EFF5"))
```

Dataset format:
```{r, echo=TRUE}
head(data, n=1)
```
NSN stands for NATO Stock Number, it's a unique identifier for military supplies across NATO nations


## Which items were purchased the most?
```{r}
itemFreq = count(data, "Item.Name")
itemFreq = itemFreq[with(itemFreq, order(-freq)), ]
head(itemFreq, n=10)
```


## Total expenditure on an item-level basis
```{r}
costs_df = itemFreq
costs_df$freq = costs_df$freq + (1 - data$Quantity[match(costs_df$Item.Name, data$Item.Name)])
costs_df$Total.Cost = data$Acquisition.Cost[match(costs_df$Item.Name, data$Item.Name)]
costs_df$Total.Cost = costs_df$Total.Cost*costs_df$freq
costs_df$NSN = data$NSN[match(costs_df$Item.Name, data$Item.Name)] # NSN will be useful to us
costs_df = costs_df[c("NSN", "Item.Name", "freq", "Total.Cost")] # reorder for readability
costs_df = costs_df[with(costs_df, order(-Total.Cost)), ]
head(costs_df, n=20)
```
\$195,868,316 spent on Humvee-like trucks, \$140,492,000 on mine-resistant trucks and $100,574,736 on helicopters were the three largest expenses.


## Extracting a purely-militarized subset
There is some equipment which is objectively non-militarized, such as the dump trucks, fork lifts and diesel generators. To focus this analysis on the militarization of US police forces we'll reduce the data to a subset that is explicitly militarized in nature, items which could serve little to no purpose in the policing of civilians. For simplicity, only the first 100 items with the highest total cost were examined manually. Other items containing keywords which indicate that they are militarized are also added.

Disclaimer: There is subjectivity as to what is "militarized" and what is not. While items were thoroughly researched by looking up their NSN, the author does not have significant knowledge of this domain.
```{r}
subset = c("TRUCK,UTILITY",  "MINE RESISTANT VEHICLE", "ONLY COMPLETE COMBAT/ASSAULT/TACTICAL WHEELED VEHICLES",
           "RIFLE,5.56 MILLIMETER", "ILLUMINATOR,INFRARED", "TRUCK,CARGO", "TRUCK,VAN",
           "CARRIER,PERSONNEL,FULL TRACKED", "TRUCK,ARMORED", "FORWARD LOOKING INFRARED IMAGING SYSTEM",
           "TRUCK,WRECKER", "NIGHT VISION GOGGLE", "VIEWER,NIGHT VISION", "BOAT,BRIDGE ERECTION,INBOARD ENGINE",
           "LIGHT ARMORED VEHICLE", "RANGE FINDER,LASER", "SIGHT,REFLEX",
           "SHOP EQUIPMENT,AUTOMOTIVE VEHICLE", "TRUCK,STAKE", "RIFLE,7.62 MILLIMETER", "MINE RESISTANT VEHI",
           "TRUCK,TANK", "SIGHT,NIGHT VISION SNIPERSCOPE", "SEMITRAILER,TANK")
subset = c(subset, unique(data$Item.Name[with(data, grepl("RIFLE", Item.Name))]))
subset = c(subset, unique(data$Item.Name[with(data, grepl("TANK", Item.Name))]))
subset = c(subset, unique(data$Item.Name[with(data, grepl("GRENADE", Item.Name))]))
subset = c(subset, unique(data$Item.Name[with(data, grepl("BAYONET", Item.Name))]))
subset = unique(subset) # remove duplicates

subset_df = data[data$Item.Name %in% subset]
```

```{r}
subset_costs_df = count(subset_df, "Item.Name")
subset_costs_df = subset_costs_df[with(subset_costs_df, order(-freq)), ]
subset_costs_df$freq = subset_costs_df$freq + (subset_df$Quantity[match(subset_costs_df$Item.Name, subset_df$Item.Name)] - 1)
subset_costs_df$Total.Cost = subset_df$Acquisition.Cost[match(subset_costs_df$Item.Name, subset_df$Item.Name)]*subset_costs_df$freq
subset_costs_df = subset_costs_df[with(subset_costs_df, order(-Total.Cost)), ]
length(subset_costs_df$Total.Cost)
sum(subset_costs_df$Total.Cost)
```
All in all, there are 175 unique items in the subset with a total cost of \$609,879,850


## State-level analysis
```{r}
# see who spent the most
suppressWarnings(
ggplot(data = head(states_df[order(-states_df$spending), ], n=10), aes(reorder(state, -spending), spending)) +
  geom_histogram(stat="identity", fill="#4b5320") + # army green
  scale_x_discrete() +
  scale_y_continuous(labels = comma, limits = c(0, 50000000),  expand=c(0,0)) +
  military_theme + theme(axis.text.x = element_text(angle=45, hjust=1)) +
  xlab("State") + ylab("Spending in USD") + ggtitle("Top 10 Spenders"))

# Adjust for population
states_pop = fread("~/projects/policeMilitarySurplus/data/statePopulation.csv")
states_df$spendingPerCapita = states_df$spending / states_pop[match(states_df$state, sapply(states_pop$state, tolower))]$pop_est_2014

suppressWarnings(
ggplot(data = head(states_df[order(-states_df$spendingPerCapita), ], n=10), aes(reorder(state, -spendingPerCapita), spendingPerCapita)) +
  geom_histogram(stat="identity", fill="#4b5320") +
  scale_x_discrete() +
  scale_y_continuous(labels = comma, limits = c(0, 10), expand=c(0,0)) +
  military_theme +
  theme(axis.text.x = element_text(angle=45, hjust=1, size = 12)) +
  xlab("State") + ylab("Spending per capita in USD") +
  ggtitle("Top 10 Spenders per Capita"))
```

```{r}
# let's see this on a map
suppressWarnings(
statebins(states_df, "stateAbr", "spending", legend_position = "right", breaks=5,
                     labels=c("<$10m","<$20m","<$30m","<$40m","<$50m"), font_size = 5, legend_title="Spending in USD",
                     brewer_pal = "Greens", text_color = "black",
                     plot_title = "Total Spending from 2006-2014",
                     title_position = "top"))
suppressWarnings(
statebins(states_df, "stateAbr", "spendingPerCapita", legend_position = "right",
                     legend_title="Spending per capita by state in USD", font_size = 5,
                     breaks = 5, labels=c("<$2","<$4","<$6","<$8","<$10"),
                     brewer_pal = "Greens", text_color = "black",
                     plot_title = "Spending per Capita from 2006-2014",
                     title_position = "top"))
```

## County-level analysis

```{r, echo=FALSE}
# Reshape data onto a county level
subset_df$County = sapply(subset_df$County, tolower)
subset_df$County = sub("^(saint |st )", "st. ", subset_df$County) # Standardize counties with Saint in the name
subset_df$State = sapply(state.name[match(subset_df$State, state.abb)], tolower)
counties_df = aggregate(subset_df$Acquisition.Cost, by=list(county=subset_df$County, state=subset_df$State), FUN=sum)
names(counties_df)[names(counties_df) == 'x'] = 'spending'
counties_df = counties_df[complete.cases(counties_df),]
counties_df = counties_df[with(counties_df, order(-spending)),]
head(counties_df, n=10)

# We'll introduce a new column to make matching easier with all of these census data sets
counties_df$Area_name = with(counties_df, paste0(stri_trans_totitle(county), ", ", sapply(state.abb[match(stri_trans_totitle(state), state.name)], toupper)))

# adjust for population 
pop_df = read_excel("~/projects/policeMilitarySurplus/data/censusCountyPopulationTotal1.xls", sheet = 1)
counties_df$population = pop_df$POP010210D[match(counties_df$Area_name, pop_df$Area_name)] # POP010210D: resident population in 2010
counties_df$spendingPerCapita = counties_df$spending / counties_df$population
counties_df = counties_df[,c('county', 'state', 'population', 'spending', 'spendingPerCapita', 'Area_name')]
counties_df = counties_df[with(counties_df, order(-spendingPerCapita)),]
counties_df$spending[counties_df$spending == 0] = NA
counties_df = na.omit(counties_df)
```

```{r}
# There are 3144 counties in the US (https://en.wikipedia.org/wiki/List_of_counties_by_U.S._state)
print(length(counties_df$county)) # Counties with spending > 0
print(length(counties_df$county)/3144)
```
Out of the 3144 counties in the US, 2378, or 75%, had some spending on militarized equipment. Counties which have spent nothing on militarized equipment are not taken into account in the following analyses.

```{r, fig.width=12, fig.height=8}
# Is population correlated with spending? 
spendPopCor = format(cor(counties_df$population, counties_df$spending, use = "complete"), digits = 4)
suppressWarnings(
ggplot(data = counties_df, aes(x = log(counties_df$population), y = log(counties_df$spending))) +
  military_theme + xlab("Population (log smoothed)") +
  ylab("Spending (log smoothed)") +
  geom_point(colour="#4b5320") + geom_smooth(method='lm', formula=y~x, colour="#454955") +
  ggtitle(paste("Correlation between spending and population = ", spendPopCor)))
```
The correlation coefficient of 0.47 implies that there is a positive relationship between a county's police departments' spending on militarized equipment and the county's population. In simpler terms, counties with more people tend to have more militarized equipment.

```{r, fig.width=12, fig.height=8}
# plot distribution of spending
suppressWarnings(
ggplot(data = counties_df, aes(x=counties_df$spending)) +
  stat_density(aes(y=..count..), color="#454955", fill="dark green", alpha=0.3) +
  stat_density(aes(y=..count..), color="#454955", fill="dark green", alpha=0.3) +
  scale_x_continuous(breaks=c(0, 100, 1000, 5000, 25000, 100000, 500000, 1000000, 3000000, 10000000),
                     labels=c("$0","$100","$1k","$5k","$25k","$100k","$500k","$1mil","$3mil","$10mil"),
                     trans="log1p", expand=c(0,0), limits=c(0,100000000)) +
  scale_y_continuous(breaks=c(0, 50, 100, 150, 200, 250), expand=c(0,0), limits=c(0,250)) +
  military_theme + xlab("Total Spending in USD") + ylab("Number of Counties") +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  ggtitle("Total Spending Density on a County Level"))
```
```{r, fig.width=12, fig.height=8}
# plot distribution of spending per capita
suppressWarnings(
ggplot(data = counties_df, aes(x=counties_df$spendingPerCapita)) +
  stat_density(aes(y=..count..), color="#454955", fill="dark green", alpha=0.3) +
  stat_density(aes(y=..count..), color="#454955", fill="dark green", alpha=0.3) +
  scale_x_continuous(breaks=c(-1,0,1,2,3,4,5,10,30,100,300,500,1000), trans="log1p", expand=c(0,0), limits=c(0,1750)) +
  scale_y_continuous(breaks=c(0,100,200,300,400,500,600,700,750), expand=c(0,0), limits=c(0,800)) +
  military_theme + xlab("Spending Per Capita in USD") + ylab("Number of counties") +
  ggtitle("Spending Per Capita Density on a County Level"))
```
```{r, fig.width=12, fig.height=8}
# plot the counties which spent the most 
suppressWarnings(
ggplot(data = highest_spenders, aes(x=reorder(Area_name, -spending), y=spending)) +
  geom_histogram(stat="identity", fill="#4b5320") +
  scale_y_continuous(breaks=c(0, 5000000, 10000000, 15000000),
                     labels=c("0", "5mil","10mil", "15mil")) +
  military_theme + theme(axis.text.x = element_text(angle=45, hjust=1), plot.title = element_text(hjust = 0)) +
  xlab("County") + ylab("Total Spending in USD") +
  ggtitle("Highest spending by county"))
```
```{r, fig.width=12, fig.height=8}
# plot the counties which spent the most per capita
suppressWarnings(
ggplot(data = head(counties_df[order(-counties_df$spendingPerCapita), ], n=10), aes(reorder(county, -spendingPerCapita), spendingPerCapita)) +
  geom_histogram(stat="identity", fill="#4b5320") +
  scale_x_discrete(labels=counties_df$Area_name) +
  scale_y_continuous(labels = comma, limits = c(0, 1250)) +
  military_theme + theme(axis.text.x = element_text(angle=45, hjust=1), plot.title = element_text(hjust = 0)) +
  xlab("County") + ylab("Spending per capita in USD") +
  ggtitle("Highest spending per capita by county"))
```




